---
title: "Vector Autoregressive Models"
subtitle: "Time Series Final | 'Flavor 2'"
author: "Marko Jurkovich & Matt Zacharski"
date: "7 June 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vars)
library(lmtest)
library(forecast)


# na 'resolver'. turns na's into previous day's close value

na2lag1 <- function(x) {
  i <- 1
  while (i <= length(x)) {
    if (is.na(x[i])) {
      x[i] <- x[i - 1]
    }
    i <- i + 1
  }
  return(x)
}
```

## Reading In Data

```{r}
# reading in data
btc <- read_csv("BTC-USD.csv")
eth <- read_csv("ETH-USD.csv")
doge <- read_csv("DOGE-USD.csv")
eth <- eth[-1827, ] # matching length
```

## Plotting the Data

```{r}
# creating closing price dataframe
closing_prices <- as.data.frame(cbind(btc$Close, eth$Close, doge$Close))
closing_prices$Date <- btc$Date
colnames(closing_prices) <- c("BTC", "ETH", "DOGE", "Date")

# plotting the 3 time series
plot(closing_prices$Date, closing_prices$BTC, type = "l")
lines(closing_prices$Date, closing_prices$ETH, type = "l", col = "red")
lines(closing_prices$Date, closing_prices$DOGE, type = "l", col = "dark green")
```

## Scaling the Data and Plotting it 

```{r}
# creating scaled data dataframe
btc_scale <- scale(closing_prices$BTC)
doge_scale <- scale(closing_prices$DOGE)
eth_scale <- scale(closing_prices$ETH)

scaled_data <- data.frame(
  Date = closing_prices$Date,
  scaled_bth = btc_scale[, 1],
  scaled_eth = eth_scale[, 1],
  scaled_doge = doge_scale[, 1]
)

# plotting the scaled data
plot(scaled_data$Date, scaled_data$scaled_bth, type = "l")
lines(scaled_data$Date, scaled_data$scaled_eth, type = "l", col = "red")
lines(scaled_data$Date, scaled_data$scaled_doge, type = "l", col = "dark green")
```

## Scaling, Logging, and Plotting the Data

```{r}

# creating scaled log dataframe
Date <- closing_prices$Date
sl_btc <- scale(log((closing_prices$BTC)))[, 1]
sl_doge <- scale(log(as.numeric(closing_prices$DOGE)))[, 1]
sl_eth <- scale(log(as.numeric(closing_prices$ETH)))[, 1]
sl_data <- data.frame(Date, sl_btc, sl_doge, sl_eth)


# plotting scaled log data
plot(sl_data$Date, sl_data$sl_btc, type = "l", main = "Plot of Scaled and Log Data", xlab = "Year", ylab = "Loged and Scaled Price")
lines(sl_data$Date, sl_data$sl_eth, type = "l", col = "red")
lines(sl_data$Date, sl_data$sl_doge, type = "l", col = "dark green")
legend("topleft", legend = c("BTC", "ETH", "DOGE"), fill = c("black", "red", "dark green"))
```

## Plotting Differenced Scaled Data 

```{r}

# differenced scaled (unlogged) data plot
plot(scaled_data$Date[c(-1, -2)], diff(scaled_data$scaled_bth, 2), type = "l")
lines(scaled_data$Date[c(-1, -2)], diff(scaled_data$scaled_eth, 2), type = "l", col = "red")
lines(scaled_data$Date[c(-1, -2)], diff(scaled_data$scaled_doge, 2), type = "l", col = "dark green")
```

## Selecting p 

```{r}

# varselect for scaled & logged data, determining optimal lag length
slvar <- VARselect(na.omit(sl_data[, -1]), lag.max = 100, type = "both")
slvar$selection
plot(slvar$criteria[1, ], type = "l")
lines(slvar$criteria[2, ], type = "l", col = "red")
```

# granger test for causality 

```{r}

grangertest(sl_btc ~ sl_doge, data = sl_data, order = 1)

grangertest(sl_btc ~ sl_eth, data = sl_data, order = 1)

grangertest(sl_doge ~ sl_eth, data = sl_data, order = 1)

grangertest(sl_doge ~ sl_btc, data = sl_data, order = 1)

grangertest(sl_eth ~ sl_btc, data = sl_data, order = 1)

grangertest(sl_eth ~ sl_doge, data = sl_data, order = 1)
```

## Fitting Models

```{r}

# removing NAs
sl_data$sl_btc <- na2lag1(sl_data$sl_btc)
sl_data$sl_doge <- na2lag1(sl_data$sl_doge)
sl_data$sl_eth <- na2lag1(sl_data$sl_eth)


# using results of VARselect to fit two models than compare them
fitvar_p9 <- VAR(sl_data[, -1], p = 9, type = "both")
fitvar_p1 <- VAR(sl_data[, -1], p = 1, type = "both")

summary(fitvar_p9)
summary(fitvar_p1)
```

## Diagnostic Tests

```{r}

# Asymptotic Portmanteau test

# we fail to find evidence of autocorrelation in the residuals of the VAR(9) model at the default of 16 lags,
serial.test(fitvar_p9, lags.pt = 16)


# we find evidence of autocorrelation in the residuals of the VAR(1) model
serial.test(fitvar_p1, lags.pt = 16)


# Breusch-Godfrey test

# same reversal present for the BG test...
serial.test(fitvar_p9, lags.bg = 16, type = "BG")


serial.test(fitvar_p1, lags.bg = 16, type = "BG")
```


```{r}
# testing the stability of the model
stablity_test_p9 <- stability(fitvar_p9, type = c("OLS-CUSUM"))
stablity_test_p1 <- stability(fitvar_p1, type = c("OLS-CUSUM"))

# the plots look decent
plot(stablity_test_p9)
plot(stablity_test_p1)
```



```{r}
# Jarque–Bera test is a goodness-of-fit test of whether sample data have the skewness and kurtosis matching a normal distribution
p9_normal <- normality.test(fitvar_p9)
p1_normal <- normality.test(fitvar_p1)

plot(p9_normal)
plot(p1_normal)


# Portmanteau Q and test for the null hypothesis that the residuals of an model are homoscedastic
arch.test(fitvar_p9)
arch.test(fitvar_p1)
```


## Residual Plots for Each Fit

```{r}

acf(residuals(fitvar_p9))


acf(residuals(fitvar_p1))
```

## Forecasting

```{r}

# need to refit model using slightly differently formatted data
sl_data_as_ts <- as.ts(sl_data[, -1])
fitvar_p9 <- VAR(sl_data_as_ts, p = 9, type = "both")

forecast(fitvar_p9, h = 12) %>%
  autoplot() + xlab("Year")

fitvar_p1 <- VAR(sl_data_as_ts, p = 1, type = "both")

forecast(fitvar_p1, h = 12) %>%
  autoplot() + xlab("Year")
```

## Evaluating Forecast Quality


```{r}

# Our data ends on 5-25-2021, so now we will load in the most recent closing prices for all three cryptocurrencies

new_btc <- read_csv("BTC-USD2.csv")
new_doge <- read_csv("DOGE-USD2.csv")
new_eth <- read_csv("ETH-USD2.csv")


closing_prices2 <- as.data.frame(cbind(new_btc$Close, new_eth$Close, new_doge$Close))
closing_prices2$Date <- new_btc$Date
colnames(closing_prices2) <- c("BTC", "ETH", "DOGE", "Date")


# creating scaled log dataframe
closing_prices_combinded <- rbind(closing_prices, closing_prices2)
Date2 <- closing_prices_combinded$Date
sl_btc2 <- scale(log((closing_prices_combinded$BTC)))[, 1]
sl_doge2 <- scale(log(as.numeric(closing_prices_combinded$DOGE)))[, 1]
sl_eth2 <- scale(log(as.numeric(closing_prices_combinded$ETH)))[, 1]
sl_data2 <- data.frame(Date2, sl_btc2, sl_doge2, sl_eth2)

# truncated the old data we used to keep the data on the same scale
sl_data2 <- sl_data2[1827:1838, ]


sl_data_as_ts2 <- as.ts(sl_data2[, -1])

# we need to get the point forecasts of our p9 model
p9_data <- forecast(fitvar_p9, h = 12)

p9_data_temp <- as.data.frame(p9_data)

p9_data_temp <- p9_data_temp[, 1:3]



# we need to change the format of our data for easier plotting
sl_data_longer <- pivot_longer(sl_data2, cols = sl_btc2:sl_eth2, names_to = "Series", values_to = "Value")
sl_data_longer$Date2 <- rep(1827:1838, times = 1, each = 3)


sl_data_longer$Time <- sl_data_longer$Date2
sl_data_longer$Date2 <- NULL


p9_data_temp$Value <- p9_data_temp$`Point Forecast`
p9_data_temp$`Point Forecast` <- NULL


all_data <- rbind(p9_data_temp, sl_data_longer)
```

```{r}
all_data %>%
  filter(Series == "sl_btc" | Series == "sl_btc2") %>%
  ggplot(aes(x = Time, y = Value, group = Series, color = Series)) +
  geom_line() +
  scale_color_discrete(
    name = "Series",
    breaks = c("sl_btc", "sl_btc2"),
    labels = c("Forecasted Data", "Actual Data")
  ) +
  labs(title = "VAR(9) Model predictions for BTC contrasted with Actual Prices")

all_data %>%
  filter(Series == "sl_doge" | Series == "sl_doge2") %>%
  ggplot(aes(x = Time, y = Value, group = Series, color = Series)) +
  geom_line() +
  scale_color_discrete(
    name = "Series",
    breaks = c("sl_doge", "sl_doge2"),
    labels = c("Forecasted Data", "Actual Data")
  ) +
  labs(title = "VAR(9) Model predictions for DOGE contrasted with Actual Prices")

all_data %>%
  filter(Series == "sl_eth" | Series == "sl_eth2") %>%
  ggplot(aes(x = Time, y = Value, group = Series, color = Series)) +
  geom_line() +
  scale_color_discrete(
    name = "Series",
    breaks = c("sl_eth", "sl_eth2"),
    labels = c("Forecasted Data", "Actual Data")
  ) +
  labs(title = "VAR(9) Model predictions for ETH contrasted with Actual Prices")
```

## Plotting VAR(1) Data and Actual Values 

```{r}
# we need to get the point forecasts of our p9 model

p1_data <- forecast(fitvar_p1, h = 12)


p1_data_temp <- as.data.frame(p1_data)

p1_data_temp <- p1_data_temp[, 1:3]



# we need to change the format of our data for easier plotting


p1_data_temp$Value <- p1_data_temp$`Point Forecast`
p1_data_temp$`Point Forecast` <- NULL


all_data2 <- rbind(p1_data_temp, sl_data_longer)
```

```{r}

all_data2 %>%
  filter(Series == "sl_btc" | Series == "sl_btc2") %>%
  ggplot(aes(x = Time, y = Value, group = Series, color = Series)) +
  geom_line() +
  scale_color_discrete(
    name = "Series",
    breaks = c("sl_btc", "sl_btc2"),
    labels = c("Forecasted Data", "Actual Data")
  ) +
  labs(title = "VAR(1) Model predictions for BTC contrasted with Actual Prices")

all_data2 %>%
  filter(Series == "sl_doge" | Series == "sl_doge2") %>%
  ggplot(aes(x = Time, y = Value, group = Series, color = Series)) +
  geom_line() +
  scale_color_discrete(
    name = "Series",
    breaks = c("sl_doge", "sl_doge2"),
    labels = c("Forecasted Data", "Actual Data")
  ) +
  labs(title = "VAR(1) Model predictions for DOGE contrasted with Actual Prices")

all_data2 %>%
  filter(Series == "sl_eth" | Series == "sl_eth2") %>%
  ggplot(aes(x = Time, y = Value, group = Series, color = Series)) +
  geom_line() +
  scale_color_discrete(
    name = "Series",
    breaks = c("sl_eth", "sl_eth2"),
    labels = c("Forecasted Data", "Actual Data")
  ) +
  labs(title = "VAR(1) Model predictions for ETH contrasted with Actual Prices")
```



## Calcualting the Percentage Differance of the Forecasts for VAR(9)

```{r}

# just loading in the above data reformatted for easier use
pct_diff_data <- read_csv("forecast_actual.csv")



btc_means <- rowMeans(pct_diff_data[, c("forecast_btc", "sl_btc2")], na.rm = TRUE)
eth_means <- rowMeans(pct_diff_data[, c("forecast_eth", "sl_eth2")], na.rm = TRUE)
doge_means <- rowMeans(pct_diff_data[, c("forecast_doge", "sl_doge2")], na.rm = TRUE)


# percentage differnce for each cryptocurrency, BTC had the lowest % differnece! So kind of looks like BTC was best predicted for the p9 model
mean((abs(pct_diff_data$forecast_btc - pct_diff_data$sl_btc2) / btc_means) * 100)

mean((abs(pct_diff_data$forecast_eth - pct_diff_data$sl_eth2) / eth_means) * 100)

mean((abs(pct_diff_data$forecast_doge - pct_diff_data$sl_doge2) / doge_means) * 100)
```
