---
title: "Vector Autoregressive Models"
author: "Marko Jurkovich & Matt Zacharski"
date: "6/3/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vars)
```

### Introduction

- Motivated by the goal of predicting a time series via a correlated partner time series, we will investigate Vector Autoregressive models (AKA VAR(p) models). 

- Why VAR models exist, what they're better at than AR models. In predictions they don't just include a variable at time t, but take all previous observations into account.

### Model

### Estimation

### Diagnostics

### Forecasting

### Data Example

### Bibliography

### Code Appendix

```{r}
btc<-read_csv('BTC-USD.csv')
eth<-read_csv('ETH-USD.csv')
doge<-read_csv('DOGE-USD.csv')
eth <- eth[-1827,]

closing_prices <- as.data.frame(cbind(btc$Close, eth$Close, doge$Close))

closing_prices$Date <- btc$Date

colnames(closing_prices) <- c('BTC', 'ETH', 'DOGE', 'Date')

plot(closing_prices$Date, closing_prices$BTC, type = 'l')
lines(closing_prices$Date, closing_prices$ETH, type = 'l', col = 'red')
lines(closing_prices$Date, closing_prices$DOGE, type = 'l', col = 'dark green')

#plot with standardized data?

btc_scale <- scale(closing_prices$BTC)
doge_scale <- scale(closing_prices$DOGE)
eth_scale <- scale(closing_prices$ETH)

scaled_data <- data.frame(Date = closing_prices$Date,
                          scaled_bth = btc_scale[,1],
           scaled_eth = eth_scale[,1],
           scaled_doge = doge_scale[,1])

plot(scaled_data$Date, scaled_data$scaled_bth, type = 'l')
lines(scaled_data$Date, scaled_data$scaled_eth, type = 'l', col = 'red')
lines(scaled_data$Date, scaled_data$scaled_doge, type = 'l', col = 'dark green')

#try to make the plots kinda stationary maybe?
plot(scaled_data$Date[c(-1,-2)], diff(scaled_data$scaled_bth, 2), type = 'l')
lines(scaled_data$Date[c(-1,-2)], diff(scaled_data$scaled_eth, 2), type = 'l', col = 'red')
lines(scaled_data$Date[c(-1,-2)], diff(scaled_data$scaled_doge, 2), type = 'l', col = 'dark green')
#eh decent I guess

scaled_data <- scaled_data[,-1]

scaled_data_clean <- na.omit(scaled_data)


fitvar = VAR(scaled_data_clean, p=2, type="both")

summary(fitvar)

acf(residuals(fitvar))


VARselect(scaled_data_clean, lag.max=12, type="const")[["selection"]]
ts_obj <- as.ts(scaled_data_clean)

var12 <- VAR(ts_obj, p=12, type="const")
serial.test(var12, lags.pt=12, type="PT.asymptotic")

library(forecast)
forecast(var12) %>%
  autoplot() + xlab("Year")

t <- as.data.frame(forecast(var12))

t$`Point Forecast` * attr(t$`Point Forecast`, 'scaled:scale') + attr(t$`Point Forecast`, 'scaled:center')


```