---
title: "Vector Autoregressive Models"
author: "Marko Jurkovich & Matt Zacharski"
date: "6/3/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

### Introduction

- Motivated by the goal of predicting a time series via a correlated partner time series, we will investigate Vector Autoregressive models (AKA VAR(p) models). 

- Why VAR models exist, what they're better at than AR models. In predictions they don't just include a variable at time t, but take all previous observations into account.

### Model

### Estimation

### Diagnostics

### Forecasting

### Data Example

### Bibliography

### Code Appendix

```{r}
btc<-read_csv('BTC-USD.csv')
eth<-read_csv('ETH-USD.csv')
doge<-read_csv('DOGE-USD.csv')
eth <- eth[-1827,]

closing_prices <- as.data.frame(cbind(btc$Close, eth$Close, doge$Close))

closing_prices$Date <- btc$Date

colnames(closing_prices) <- c('BTC', 'ETH', 'DOGE', 'Date')

plot(closing_prices$Date, closing_prices$BTC, type = 'l')
lines(closing_prices$Date, closing_prices$ETH, type = 'l', col = 'red')
lines(closing_prices$Date, closing_prices$DOGE, type = 'l', col = 'dark green')

#plot with standardized data?



```